% future/fp.tex
% mainfile: ../perfbook.tex
% SPDX-License-Identifier: CC-BY-SA-3.0

\section{Functional Programming for Parallelism}
\label{sec:future:Functional Programming for Parallelism}
%
\epigraph{The curious failure of functional programming for parallel
	  applications.}
	 {\emph{Malte Skarupke}}

1980년대 초에 제가 처음으로 함수형 프로그래밍 수업을 들었을 때, 교수님은
부작용으로부터 자유로운 함수형 프로그래밍 스타일은 사소한 병렬화와 분석에 잘
맞는다고 단정하셨습니다.
30년 후, 이 단정은 남아있습니다만, 주류 제작사에서의 병렬 함수형 언어의 사용은
매우 적으며, 이는 교수님의 프로그램은 상태를 유지하지 않고 I/O 를 하지 않아야
한다고 하셨던 추가적인 말씀과 관계가 아주 없지는 않을 겁니다.
Erlang 과 같은 함수형 언어의 틈색에서의 사용이 있고, 여러 다른 함수형 언어에
멀티쓰레드 지원이 추가되었습니다만, 주류 제조 환경은 C, C++, Java, 그리고
Fortran 같은 절차형 언어의 (보통 OpenMP, MPI, 또는 coarrays 와 증간된 형태의)
영역으로 남아 있습니다.

이 상황은 자연스레 ``분석이 목표라면, 분석 전에 왜 이 절차형 언어를 함수형
언어로 변환하면 어떨까?'' 라는 질문을 일으키게 합니다.
이 방법에 대한 여러 반대들이 물론 존재하는데, 이 중 세개만 나열하자면:

\iffalse

When I took my first-ever functional-programming class in the early 1980s,
the professor asserted that the side-effect-free functional-programming
style was well-suited to trivial parallelization and analysis.
Thirty years later, this assertion remains, but mainstream production
use of parallel functional languages is minimal, a state of affairs that
might not be entirely unrelated to professor's additional
assertion that programs should neither maintain state nor do I/O\@.
There is niche use of functional languages such as Erlang, and
multithreaded support has been added to several other functional languages,
but mainstream production usage remains the province of procedural
languages such as C, C++, Java, and Fortran (usually augmented with
OpenMP, MPI, or coarrays).

This situation naturally leads to the question ``If analysis is the goal,
why not transform the procedural language into a functional language before
doing the analysis?''
There are of course a number of objections to this approach, of which
I list but three:

\fi

\begin{enumerate}
\item	절차형 언어는 다른 독립된 함수들에 의해, 또는 더 나쁘게도 여러 쓰레드에
	의해 수정될 수 있는 전역 변수들을 종종 많이 사용합니다.
	Haskell 의 \emph{monads} 는 싱글 쓰레드 전역 상태를 처리하기 위해
	발명되었으며, 전역 상태로의 멀티 쓰레드에 의한 액세스는 함수형 모델에
	추가적인 위반을 가함을 알아 두시기 바랍니다.
\item	멀티쓰레드 기반 절차형 언어는 종종 락, 어토믹 오퍼레이션, 그리고
	트랜잭션 같은 동기화 기능을 사용하는데, 이는 함수형 모델에 위반을
	더합니다.
\item	절차형 언어는 함수 인자에 \emph{별명을 붙일 (alias)} 수 있는데, 예를
	들어 같은 구조체로의 포인터를 어떤 함수의 같은 호출에 두개의 다른
	인자로 넘길 수 있습니다.
	이는 이 함수가 이 구조체를 두개의 다른 (그리고 아마도 겹치는) 코드
	순서에서 알지 못한 업데이트를 만들 수 있게 하는데, 이는 분석을 무척
	복잡하게 만듭니다.

\iffalse

\item	Procedural languages often make heavy use of global variables,
	which can be updated independently by different
	functions, or, worse yet, by multiple threads.
	Note that Haskell's \emph{monads} were invented to deal with
	single-threaded global state, and that multi-threaded access to
	global state inflicts additional violence on the functional model.
\item	Multithreaded procedural languages often use synchronization
	primitives such as locks, atomic operations, and transactions,
	which inflict added violence upon the functional model.
\item	Procedural languages can \emph{alias} function arguments,
	for example, by passing a pointer to the same structure via two
	different arguments to the same invocation of a given function.
	This can result in the function unknowingly updating that
	structure via two different (and possibly overlapping) code
	sequences, which greatly complicates analysis.

\fi

\end{enumerate}

물론, 전역 상태, 동기화 기능, 그리고 aliasing 의 중요성을 감안하여 현명한
함수형 프로그래밍 전문가들은 함수형 프로그래밍 모델을 그것들에 조정하기 위한
여러 시도를 제안했으면, monads 그런 것들 중 하나입니다.

또다른 방법은 병렬 절차적 프로그램을 함수형 그로그램으로 변환하고, 그 결과를
분석하기 위해 함수형 프로그래밍 도구를 사용하는 것입니다.
하지만 이것보다 훨씬 더 잘할 수 있는데, 모든 실제 계산은 유한한 입력을 갖는
거대한 유한 상태 기계 (finite-state machine) 이기 때문입니다.
이는 모든 실제 프로그램은 하나의 표현으로 변환될 수 있다는 것을 의미하는데,
비실용적으로 큰 것일 수도 있긴 합니다~\cite{VijayDSilva2012-sas}.

\iffalse

Of course, given the importance of global state, synchronization
primitives, and aliasing, clever functional-programming experts have
proposed any number of attempts to reconcile the function programming
model to them, monads being but one case in point.

Another approach is to compile the parallel procedural program into
a functional program, then to use functional-programming tools to analyze
the result.
But it is possible to do much better than this, given that any real
computation is a large finite-state machine with finite input that
runs for a finite time interval.
This means that any real program can be transformed into an expression,
possibly albeit an impractically large one~\cite{VijayDSilva2012-sas}.

\fi

하지만, 병렬 알고리즘의 여러 저수준 핵심 부분들이 현대의 컴퓨터의 메모리에 쉽게
들어갈 수 있을 만큼 충분히 작은 표현으로 변환됩니다.
만약 그런 표현이 단정과 결합된다면, 그 단정이 격발될 수 있을지 확인하는 것은
satisfiability 문제가 됩니다.
Satisfiability 문제는 NP-complete 하긴 하나, 전체 상태 공간을 생성하는데 필요한
것보다 훨씬 짧은 시간에 종종 해결될 수 있습니다.
또한, 이 해결 시간은 기저의 메모리 모델에 덜 종속적인 것으로 보이므로, 완화된
순서 규칙의 시스템에서 돌아가는 알고리즘도 검사될 수
있습니다~\cite{JadeAlglave2013-cav}.

일반적인 방법은 프로그램을 single-static-assignment (SSA) 형태로 변환하여서 각
변수로의 할당문이 그 변수의 분리된 버전을 만들게 하는 겁니다.
이는 모든 활성화된 쓰레드로부터의 할당문에 적용되어 그 결과 만들어지는 표현은
해당 코드의 모든 가능한 수행을 갖게 됩니다.
할당문을 하나 추가하는 것은 입력과 초기 값의 어떤 조합이 어떤 단정문을
격발시키는 결과를 낳을 수 있는지 질문하게 만드는데, 이는 앞에서도 언급된 것처럼
satisfiability 문제입니다.

\iffalse

However, a number of the low-level kernels of parallel algorithms transform
into expressions that are small enough to fit easily into the memories
of modern computers.
If such an expression is coupled with an assertion, checking to see if
the assertion would ever fire becomes a satisfiability problem.
Even though satisfiability problems are NP-complete, they can often
be solved in much less time than would be required to generate the
full state space.
In addition, the solution time appears to be only weakly dependent on
the underlying memory model, so that algorithms running on weakly ordered
systems can also be checked~\cite{JadeAlglave2013-cav}.

The general approach is to transform the program into single-static-assignment
(SSA) form, so that each assignment to a variable creates a separate
version of that variable.
This applies to assignments from all the active threads, so that the
resulting expression embodies all possible executions of the code
in question.
The addition of an assertion entails asking whether any combination of
inputs and initial values can result in the assertion firing, which,
as noted above, is exactly the satisfiability problem.

\fi

한가지 가능한 반대는 이게 임의의 반복문을 우아하게 처리하지 않는다는 겁니다.
그러나, 많은 경우 이는 그 반복문을 유한한 횟수만큼 unrolling 함으로써 처리될 수
있습니다.
또한, 어떤 반복문은 귀납법으로 풀어질 수 있는 것으로 증명될 겁니다.

또다른 가능한 반대는 spinlock 이 임의의 긴 반복문을 갖고, 모든 유한 횟수
unrolling 은 이 spinlock 의 전체 행동을 포착하지 못할 거라는겁니다.
이 반대는 쉽게 극복될 수 있는 것으로 드러났습니다.
전체 spinlock 을 모델링하는 대신, 그 락을 획득하려 시도하는 trylock 을
모델링하고 즉각 그러지 못한다면 중단하는 겁니다.
그러면 이 단정문은 spinlock 이 그 락을 즉각 얻을 수 없기 때문에 중단되었을
때에는 격발되지 않도록 조정되어야만 합니다.
이 논리 표현은 시간에 무관하므로, 모든 가능한 동시성 동작은 이 방법으로 포착될
겁니다.

\iffalse

One possible objection is that it does not gracefully handle arbitrary
looping constructs.
However, in many cases, this can be handled by unrolling the loop a
finite number of times.
In addition, perhaps some loops will also prove amenable to collapse
via inductive methods.

Another possible objection is that spinlocks involve arbitrarily long
loops, and any finite unrolling would fail to capture the full behavior
of the spinlock.
It turns out that this objection is easily overcome.
Instead of modeling a full spinlock, model a trylock that attempts to
obtain the lock, and aborts if it fails to immediately do so.
The assertion must then be crafted so as to avoid firing in cases
where a spinlock aborted due to the lock not being immediately available.
Because the logic expression is independent of time, all possible
concurrency behaviors will be captured via this approach.

\fi

마지막 반대는 이 기법이 리눅스 커널을 만드는 수백만 행의 코드와 같은 완전한
크기의 소프트웨어 작품은 처리하지 못할 거라는 겁니다.
이는 그럴 수 있는 이야기이나, 리눅스 커널의 많은 작은 병렬 기능들 각각을 세세히
검증할 수 있다는 것은 상당히 가치있을 겁니다.
그리고 이 방법의 선봉에 선 연구자들은 실제로 리눅스 커널의 Tree RCU 구현을
포함한 실제 세계의 사소하지 않은 코드에
적용했습니다~\cite{LihaoLiang2016VerifyTreeRCU,MichalisKokologiannakis2017NidhuggRCU}.

이 기법이 얼마나 널리 적용가능할지 두고 봐야겠습니다만, 이는 정형 검증
영역에서의 흥미로운 혁신들 중 하나일 뿐입니다.
병렬 프로그래밍 대변자들은 그들의 피할 수 없는 함수형 프로그래밍의 세계 정복에
대한 단정에 대해 긴 시간 후에는 옳았음으로 드러날 수도 있겠으나, 이 오랜 시간
있어온 방법론은 그것의 정형 검증 영역에서는 믿을 수 있는 결과를 보이기 시작하고
있습니다.
따라서 함수형 프로그래밍의 피할 수 없는 세계 정복에 대한 의심을 계속할 이유가
있습니다.

\iffalse

A final objection is that this technique is unlikely to be able to handle
a full-sized software artifact such as the millions of lines of code making
up the Linux kernel.
This is likely the case, but the fact remains that exhaustive validation
of each of the much smaller parallel primitives within the Linux kernel
would be quite valuable.
And in fact the researchers spearheading this approach have applied it
to non-trivial real-world code, including the Tree RCU implementation in
the Linux
kernel~\cite{LihaoLiang2016VerifyTreeRCU,MichalisKokologiannakis2017NidhuggRCU}.

It remains to be seen how widely applicable this technique is, but
it is one of the more interesting innovations in the field of
formal verification.
Although it might well be that the functional-programming advocates
are at long last correct in their assertion of the inevitable
dominance of functional programming, it is clearly the case
that this long-touted methodology is starting to see credible
competition on its formal-verification home turf.
There is therefore continued reason to doubt the inevitability of
functional-programming dominance.

\fi
