% together/count.tex
% mainfile: ../perfbook.tex
% SPDX-License-Identifier: CC-BY-SA-3.0

\section{Counter Conundrums}
\label{sec:together:Counter Conundrums}
%
\epigraph{Ford carried on counting quietly.
	  This is about the most aggressive thing you can do to a
	  computer, the equivalent of going up to a human being and saying
	  ``Blood \dots blood \dots blood \dots blood \dots''}
	 {\emph{Douglas Adams}}

이 \lcnamecref{sec:together:Counter Conundrums} 는 카운터 수수께끼들에 대한
해결책들을 정리해 봅니다.

\iffalse

This \lcnamecref{sec:together:Counter Conundrums}
outlines solutions to counter conundrums.

\fi

\subsection{Counting Updates}
\label{sec:together:Counting Updates}

Schr\"odinger
(\cref{sec:datastruct:Motivating Application} 참고) 가 각 동물의 업데이트
횟수를 세고 싶어하며, 이 업데이트는 데이터 원소별 락을 이용해 동기화 된다고
해봅시다.
이 카운팅은 어떻게 해야 가장 잘 할 수 있을까요?

물론, \cref{chp:Counting} 에서의 카운팅 알고리즘들이 얼마든지 충분할 수도
있지만, 최적의 방법은 상당히 간단합니다.
각 데이터 원소에 카운터를 놓고, 그 원소의 락의 보호 아래 그 카운터를 증가시키는
겁니다!

읽기 쓰레드가 락 없이 이 숫자에 접근하면, 업데이트 쓰레드는 이 카운터를
업데이트 하는데에 \co{WRITE_ONCE()} 를 사용하고 락 없는 읽기 쓰레드는 이를 읽기
위해 \co{READ_ONCE()} 를 사용해야 합니다.

\iffalse

Suppose that Schr\"odinger (see
\cref{sec:datastruct:Motivating Application})
wants to count the number of updates for each animal,
and that these updates are synchronized using a per-data-element lock.
How can this counting best be done?

Of course, any number of counting algorithms from \cref{chp:Counting}
might qualify, but the optimal approach is quite simple.
Just place a counter in each data element, and increment it under the
protection of that element's lock!

If readers access the count locklessly, then updaters should use
\co{WRITE_ONCE()} to update the counter and lockless readers should
use \co{READ_ONCE()} to load it.

\fi

\subsection{Counting Lookups}
\label{sec:together:Counting Lookups}

Schr\"odinger 는 각 동물을 위한 탐색 횟수도 세고 싶어하며, 탐색은 RCU 로
보호된다고 해봅시다.
이 경우에 무엇이 최선의 카운팅 방법일까요?

한가지 방법은
\cref{sec:together:Counting Updates} 에서 이야기 되었듯 이 탐색 카운터를 원소별
락을 이용해 보호하는 것입니다.
불행히도, 이는 각 탐색이 락을 획득할 것을 필요로 해서 큰 시스템에서는 상당한
병목지점이 될 겁니다.

또다른 방법은 카운팅이 ``안된다고 말하기'' 로, \co{noatime} mount 옵션의 예를
따릅니다.
이 방법이 적용 가능하다면, 이게 분명 최선입니다:  어쨌건, 아무것도 안하는
것보다 빠른건 없습니다.
이 탐색 카운트가 없어질 수 없다면, 계속 읽으세요!

\iffalse

Suppose that Schr\"odinger also wants to count the number of lookups for
each animal, where lookups are protected by RCU\@.
How can this counting best be done?

One approach would be to protect a lookup counter with the per-element
lock, as discussed in \cref{sec:together:Counting Updates}.
Unfortunately, this would require all lookups to acquire this lock,
which would be a severe bottleneck on large systems.

Another approach is to ``just say no'' to counting, following the example
of the \co{noatime} mount option.
If this approach is feasible, it is clearly the best:  After all, nothing
is faster than doing nothing.
If the lookup count cannot be dispensed with, read on!

\fi

\Cref{chp:Counting} 에서의 카운터들이 무엇이든 사용될 수 있겠는데,
\cref{sec:count:Statistical Counters} 의 통계적 카운터가 아마 가장 흔한 선택일
겁니다.
하지만, 이는 큰 메모리 사용량을 초래합니다: 필요한 카운터의 수는 데이터 원소의
수 곱하기 쓰레드의 수입니다.

이 메모리 오버헤드가 지나치다면,
\cref{fig:datastruct:Read-Only Hash-Table Performance For Schroedinger's Zoo; 448 CPUs}
에 보인 해쉬 테이블 성능을 참고해 CPU 별 카운터 대신 코어별 또는 소켓별
카운터를 두는 것이 한 방편이 되겠습니다.
이는 이 카운터 값 증가가 어토믹 오퍼레이션이 될 것을 필요로 하는데, 특정
쓰레드가 언제든 다른 CPU 로 옮겨질 수 있는 사용자 모드 수행에서는 특히
그렇습니다.

어떤 원소들이 매우 빈번하게 탐색된다면, 쓰레드별로 특정 원소를 위한 로그
항목들이 병합될 수 있는 로그를 두어서 업데이트를 몰아서 하는 방법이 여럿
있습니다.
특정 로그 항목이 충분히 큰 값 증가를 가지거나 충분한 시간이 흐른 후에는 이 로그
항목들이 연관된 데이터 원소에 적용될 수 있습니다.
Silas Boyd-Wickizer 는 이 방법들을 정형화
시켰습니다~\cite{SilasBoydWickizerPhD}.

\iffalse

Any of the counters from \cref{chp:Counting}
could be pressed into service, with the statistical counters described in
\cref{sec:count:Statistical Counters} being perhaps the most common choice.
However, this results in a large memory footprint: The number of counters
required is the number of data elements multiplied by the number of
threads.

If this memory overhead is excessive, then one approach is to keep
per-core or even per-socket counters rather than per-CPU counters,
with an eye to the hash-table performance results depicted in
\cref{fig:datastruct:Read-Only Hash-Table Performance For Schroedinger's Zoo; 448 CPUs}.
This will require that the counter increments be atomic operations,
especially for user-mode execution where a given thread could migrate
to another CPU at any time.

If some elements are looked up very frequently, there are a number
of approaches that batch updates by maintaining a per-thread log,
where multiple log entries for a given element can be merged.
After a given log entry has a sufficiently large increment or after
sufficient time has passed, the log entries may be applied to the
corresponding data elements.
Silas Boyd-Wickizer has done some work formalizing this
notion~\cite{SilasBoydWickizerPhD}.

\fi
