% appendix/questions/questions.tex
% mainfile: ../../perfbook.tex
% SPDX-License-Identifier: CC-BY-SA-3.0

\QuickQuizChapter{cha:app:Important Questions}{Important Questions}{qqzquestions}
%
\Epigraph{Ask me no questions, and I'll tell you no fibs.}
	 {\emph{``She Stoops to Conquer'', Oliver Goldsmith}}

다음 섹션들은 SMP 프로그래밍에 연관된 일부 중요한 질문들을 논합니다.
각 섹션은 또한 연관된 질문에 대해 걱정하는 걸 어떻게 막는지 보이는데, 여러분의
목표가 단지 여러분의 SMP 코드가 가능한한 빨리 고통 없이 동작하게 하는
것이라면---그나저나, 이는 훌륭한 목표죠!--- 무척 중요할 수 있습니다.

이 질문들에 대한 답은 종종 단일 쓰레드 환경에서의 것에 비해 덜 직관적이지만,
약간의 노력을 한다면 이해하기 그렇게 어렵지 않습니다.
여러분이 재귀를 이해했다면, 여기에 엄청난 도전을 일으키는 건 없습니다.

\iffalse

The following sections discuss some important questions relating to
SMP programming.
Each section also shows how to avoid worrying about
the corresponding question, which can be extremely important if
your goal is to simply get your SMP code working as quickly and
painlessly as possible---which is an excellent goal, by the way!

Although the answers to these questions are often less
intuitive than they would be in a single-threaded setting,
with a bit of work, they are not that difficult to understand.
If you managed to master recursion, there is nothing here that should
pose an overwhelming challenge.

\fi

% @@@ roadmap...

% @@@ doesn't parallel automatically make things faster? (locked increment)
% @@@ doesn't getting rid of locks get rid of all these problems? (atomic inc)
% @@@ doesn't NBS take care of all these problems? (cite IPDPS paper)
% @@@ why isn't more software correct?

\input{appendix/questions/after}
\input{appendix/questions/concurrentparallel}
\input{appendix/questions/time}
\input{appendix/questions/ordering}

\QuickQuizAnswersChp{qqzquestions}
