% appendix/rcuimpl/rcu.tex
% SPDX-License-Identifier: CC-BY-SA-3.0

\QuickQuizChapter{app:rcuimpl:Read-Copy Update Implementations}{Read-Copy Update Implementations}

이 부록은 완전한 기능을 갖춘 제품 수준의 RCU 구현들 몇가지를 설명합니다.
이 구현들에 대한 이해를 위해서는
Chapter~\ref{chp:Introduction} 와
\ref{chp:Deferred Processing} 에 대한 깊은 이해는 물론이고 리눅스 커널에 대한
충분한 이해를 필요로 하는데, 뒤의 것은 여러 책들과 웹사이트를 통해 얻을 수 있을
겁니다~\cite{BovetCesati2005,CorbetRubiniKroahHartman,CorbetLWN,RobertLove2005}.

RCU 구현들을 처음 접한다면, 아마도
Appendix~\ref{chp:app:``Toy'' RCU Implementations} 에서 찾을 수 있을, 더
간단한 ``toy'' RCU 구현부터 시작해야 할겁니다.
\iffalse

This appendix describes several fully functional production-quality RCU
implementations.
Understanding of these implementations requires a thorough understanding
of the material in
Chapters~\ref{chp:Introduction} and
\ref{chp:Deferred Processing},
as well as a reasonably good understanding of the Linux kernel,
the latter of which may be found in several textbooks and
websites~\cite{BovetCesati2005,CorbetRubiniKroahHartman,CorbetLWN,RobertLove2005}.

If you are new to RCU implementations, you should start with the
simpler ``toy'' RCU implementations that may be found in
Appendix~\ref{chp:app:``Toy'' RCU Implementations}.
\fi

Section~\ref{app:rcuimpl:Sleepable RCU Implementation} 에서는 ``Sleepable RCU''
또는 SRCU 라 불리는 것을 설명하는데, 이는 SRCU 읽기 쓰레드들이 임의적으로 sleep
상태에 빠지는 것을 가능하게 합니다.
이는 제품 수준의 RCU 구현들이 그렇듯이 간단한 구현이며, 그런 구현들에 대해
공부를 시작하기에 좋은 부분입니다.

Section~\ref{app:rcuimpl:rcutree:Hierarchical RCU Overview} 은 수천개의 CPU
들을 갖춘 SMP 시스템을 위해 설계된, Classic RCU 의 높은 확장성을 갖춘 구현에
대한 개괄을 제공합니다.
Section~\ref{app:rcuimpl:rcutreewt:Hierarchical RCU Code Walkthrough} 는 이
구현에 대한 (2008년 경의) 코드의 walkthrough 를 취해봅니다.

마지막으로,
Section~\ref{app:rcuimpl:Preemptible RCU}
에서는 real-time 시스템에서 사용되는 preemptible RCU 구현에 대해 자세히
살펴봅니다.
\iffalse

Section~\ref{app:rcuimpl:Sleepable RCU Implementation} presents
``Sleepable RCU'', or SRCU, which allows SRCU readers to sleep
arbitrarily.
This is a simple implementation, as production-quality RCU implementations
go, and a good place to start learning about such implementations.

Section~\ref{app:rcuimpl:rcutree:Hierarchical RCU Overview}
gives an overview of a highly scalable implementation of Classic
RCU, designed for SMP systems sporting thousands of CPUs.
Section~\ref{app:rcuimpl:rcutreewt:Hierarchical RCU Code Walkthrough}
takes the reader on a code walkthrough of this same implementation
(as of late 2008).

Finally,
Section~\ref{app:rcuimpl:Preemptible RCU}
provides a detailed view of the preemptible RCU implementation used
in real-time systems.
\fi

% RCU requirements and desiderata?  Or at end?
\input{appendix/rcuimpl/srcu.tex}
% \input{appendix/rcuimpl/rcuclassic.tex}  @@@ from Ph.D. dissertation.
\input{appendix/rcuimpl/rcutree.tex}
\input{appendix/rcuimpl/rcutreewt.tex}
\input{appendix/rcuimpl/rcupreempt.tex}
