% defer/rcu.tex
% mainfile: ../perfbook.tex
% SPDX-License-Identifier: CC-BY-SA-3.0

\section{Read-Copy Update (RCU)}
\label{sec:defer:Read-Copy Update (RCU)}
%
\epigraph{``Free'' is a \emph{very} good price!}{\emph{Tom Peterson}}

앞의 섹션들에서 다루어진 모든 메커니즘은 특정한 행동을 그것이 안전히 이루어질
수 있을 때까지 미루는 여러 전략 중 하나를 사용했습니다.
Section~\ref{sec:defer:Reference Counting}
에서 다룬 레퍼런스 카운터는 읽기 쓰레드를 방해할 수 있는 행동을 뒤로 미루기
위해 명시적 카운터를 사용했고 이는 읽기 쪽 컨텐션과 그로 인한 낮은 확장성의
결과를 가져왔습니다.
Section~\ref{sec:defer:Hazard Pointers}
은 쓰레드별 포인터 리스트의 모습을 한 암시적 카운터를 사용했습니다.
이는 읽기 쪽 컨텐션은 제거했으나, 읽기 쓰레드가 쓰기와 조건적 브랜치는 물론,
읽기쪽 완전한 메모리 배리어 또는 리억타임에 친화적이지 않은 프로세서간 인터럽트
(inter-processor interrupt) 기능을 사용해야 했습니다.\footnote{
	어떤 중요한 특수 경우에는, 이 추가적 일이 Folly 오픈소스 라이브러리
	(\url{https://github.com/facebook/folly}) 에 구현되어 있는
	UnboundedQueue 와 ConcurrentHashMap 데이터 구조에서 보여지듯 연결
	카운팅을 사용하는 것으로 제거될 수 있습니다.}
Section~\ref{sec:defer:Sequence Locks}
에서 선보인 시퀀스 락 또한 읽기 쪽 컨텐션은 제거하나, 포인터 순회를 보호하지
않으며, 해저드 포인터와 같이, 읽기 쪽에서의 완전한 메모리 배리어 또는 업데이트
쪽에서의 프로세서간 인터럽트를 필요로 합니다.
이 방법들의 단점들은 더 나은 것은 불가능한지 질문을 던지게 합니다.

\iffalse

All of the mechanisms discussed in the preceding sections
used one of a number of approaches to defer specific actions
until they may be carried out safely.
The reference counters discussed in
Section~\ref{sec:defer:Reference Counting}
use explicit counters to defer actions that could disturb readers,
which results in read-side contention and thus poor scalability.
The hazard pointers covered by
Section~\ref{sec:defer:Hazard Pointers}
uses implicit counters in the guise of per-thread lists of pointer.
This avoids read-side contention, but requires readers to do stores and
conditional branches, as well as either full memory barriers in read-side
primitives or real-time-unfriendly inter-processor interrupts in
update-side primitives.\footnote{
	In some important special cases, this extra work can be avoided
	by using link counting as exemplified by the UnboundedQueue
	and ConcurrentHashMap data structures implemented in Folly
	open-source library
	(\url{https://github.com/facebook/folly}).}
The sequence lock presented in
Section~\ref{sec:defer:Sequence Locks}
also avoids read-side contention, but does not protect pointer
traversals and, like hazard pointers, requires either full memory barriers
in read-side primitives, or inter-processor interrupts in update-side
primitives.
These schemes' shortcomings raise the question of
whether it is possible to do better.

\fi

이 섹션은 \emph{read-copy update} (RCU) 를 소개하는데, 자주 업데이트 되는 공유
데이터로의 비싼 쓰기 없이 읽기 쓰레드가 소스 코드의 특정 영역과 연관지어지게
하는 API 를 제공합니다.
이 섹션의 나머지 부분은 여러 다른 관점에서 RCU 를 알아봅니다.
\Cref{sec:defer:Introduction to RCU} 는 고전적 RCU 소개를 제공하고,
\cref{sec:defer:RCU Fundamentals} 는 기본적인 RCU 컨셉을 다루며,
\cref{sec:defer:RCU Linux-Kernel API} 는 RCU 의 리눅스 커널 API 를 선보이고,
\cref{sec:defer:RCU Usage} 는 흔한 RCU 사용 예를 소개하고,
\cref{sec:defer:RCU Related Work} 는 RCU 와 연관된 최근의 작업들을 다루고,
\cref{sec:defer:RCU Exercises} 는 일부 RCU 연습문제를 제공하고, 마지막으로
\cref{sec:defer:What About Updates?} 는 업데이트를 다룹니다.

\iffalse

This section introduces \emph{read-copy update} (RCU), which provides
an API that allows readers to be associated with regions in the source code,
rather than with expensive updates to frequently updated shared data.
The remainder of this
section examines RCU from a number of different perspectives.
\Cref{sec:defer:Introduction to RCU} provides the classic
introduction to RCU,
\cref{sec:defer:RCU Fundamentals} covers fundamental RCU
concepts,
\cref{sec:defer:RCU Linux-Kernel API} presents the Linux-kernel
API,
\cref{sec:defer:RCU Usage} introduces some common RCU use cases,
\cref{sec:defer:RCU Related Work} covers recent work related
to RCU,
\cref{sec:defer:RCU Exercises} provides some RCU exercises,
and finally
\cref{sec:defer:What About Updates?}
discusses updates.

\fi

\input{defer/rcuintro}
\input{defer/rcufundamental}
\input{defer/rcuapi}
\input{defer/rcuusage}
\input{defer/rcurelated}
